---
title: "Getting Started with GlobalTechAnalytics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with GlobalTechAnalytics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(GlobalTechAnalytics)
library(dplyr)
library(ggplot2)
```

## Introduction

GlobalTechAnalytics is a comprehensive R package for workforce analytics and predictive modeling.
This document will walk you through the basic functionality and get you started with analyzing your organization's talent data.

## Quick Start

### Generate Sample Data

The package includes a data generation system that creates employee and project data for demonstration and testing purposes.

```{r}
# Generate sample data for demonstration
sample_data <- generate_globaltech_data(n_employees = 500, n_projects = 25)

# Explore the employee data structure
str(sample_data$employees[1:5, 1:8])  # First 5 rows, 8 columns
```

### Key Features Overview

```{r}
# Summary of generated data
cat("=== Sample Data Summary ===\n")
cat("Employees:", nrow(sample_data$employees), "\n")
cat("Projects:", nrow(sample_data$projects), "\n")
cat("Departments:", length(unique(sample_data$employees$department)), "\n")
cat("Attrition Rate:", round(mean(sample_data$employees$left_company) * 100, 1), "%\n")
```

## Interactive Dashboard

### Launch the Dashboard

The easiest way to explore your data is through the interactive Shiny dashboard:

```{r, eval=FALSE}
# Launch the interactive dashboard
launch_dashboard()
```

This opens a web-based dashboard with: - **Executive Overview**: Key metrics and strategic insights - **Attrition Analysis**: Risk prediction and employee segmentation - **Resource Optimization**: Utilization analysis and capacity planning - **Project Analytics**: Delivery performance and success factors

## Core Analytics Functions

### 1. Attrition Prediction

Build machine learning models to predict which employees are at risk of leaving:

```{r}
# Build attrition prediction model
attrition_model <- build_attrition_model(sample_data$employees)

# View model performance
cat("Model AUC:", round(attrition_model$auc, 3), "\n")

# View top risk factors
head(attrition_model$feature_importance, 8)
```

### 2. Predict Individual Risk

Use the trained model to assess specific employees:

```{r}
# Get predictions for high-risk employees
#high_risk_employees <- sample_data$employees %>%
#  filter(!left_company) %>%
#  slice_head(n = 10)

#risk_predictions <- predict_attrition_risk(attrition_model, high_risk_employees)
#head(risk_predictions)
```

### 3. Resource Utilization Analysis

Analyze how effectively your organization uses its human resources:

```{r}
# Analyze resource utilization
utilization_analysis <- analyze_resource_utilization(
  sample_data$employees, 
  sample_data$projects
)

# View department efficiency
utilization_analysis$department_utilization
```

### 4. Business Insights

Generate executive-level insights and recommendations:

```{r}
# Generate comprehensive business insights
insights <- generate_business_insights(
  sample_data$employees, 
  sample_data$projects
)

# View key metrics
insights$key_metrics
```

```{r}
# View executive summary
cat(insights$executive_summary)
```

## Data Visualization Examples

### Attrition Risk by Department

```{r}
# Create risk analysis visualization
risk_by_dept <- sample_data$employees %>%
  filter(!left_company) %>%
  group_by(department) %>%
  summarise(
    employees = n(),
    avg_risk = mean(attrition_risk_score),
    high_risk_count = sum(attrition_risk_score > 0.7),
    .groups = "drop"
  )

ggplot(risk_by_dept, aes(x = reorder(department, avg_risk), y = avg_risk)) +
  geom_col(fill = "steelblue", alpha = 0.7) +
  geom_text(aes(label = high_risk_count), hjust = -0.1) +
  coord_flip() +
  labs(
    title = "Average Attrition Risk by Department",
    x = "Department",
    y = "Average Risk Score",
    caption = "Numbers show count of high-risk employees"
  ) +
  theme_minimal()
```

### Performance vs Satisfaction Analysis

```{r}
# Analyze the relationship between performance and satisfaction
perf_sat_data <- sample_data$employees %>%
  filter(!left_company) %>%
  mutate(
    risk_level = case_when(
      attrition_risk_score >= 0.7 ~ "High Risk",
      attrition_risk_score >= 0.4 ~ "Medium Risk",
      TRUE ~ "Low Risk"
    )
  )

ggplot(perf_sat_data, aes(x = job_satisfaction, y = performance_rating, color = risk_level)) +
  geom_point(alpha = 0.6, size = 2) +
  scale_color_manual(values = c("Low Risk" = "green", "Medium Risk" = "orange", "High Risk" = "red")) +
  labs(
    title = "Performance vs Satisfaction by Risk Level",
    x = "Job Satisfaction",
    y = "Performance Rating",
    color = "Attrition Risk"
  ) +
  theme_minimal()
```

## Working with Your Own Data

### Required Data Format

To use GlobalTechAnalytics with your own data, ensure your datasets include these key columns:

#### Employee Data

-   `employee_id`: Unique identifier
-   `department`: Employee's department
-   `tenure_years`: Years with company
-   `job_satisfaction`: Rating (1-5 scale)
-   `performance_rating`: Rating (1-5 scale)
-   `age`: Employee age
-   `base_salary`: Annual salary
-   `left_company`: Boolean (TRUE if employee has left)

#### Project Data

-   `project_id`: Unique identifier
-   `project_type`: Type of project
-   `budget`: Project budget
-   `start_date`: Project start date
-   `planned_end_date`: Planned completion date
-   `status`: Current status (Active, Completed, etc.)

### Loading Your Data

```{r, eval=FALSE}
# Load your own data
your_employees <- read.csv("your_employee_data.csv")
your_projects <- read.csv("your_project_data.csv")

# Validate data format
validate_dataframe(your_employees, 
                  c("employee_id", "department", "job_satisfaction"), 
                  min_rows = 100)

# Run analytics on your data
your_model <- build_attrition_model(your_employees)
your_insights <- generate_business_insights(your_employees, your_projects)
```

## Advanced Features

### Custom Risk Thresholds

```{r}
# Analyze different risk thresholds
risk_analysis <- sample_data$employees %>%
  filter(!left_company) %>%
  summarise(
    high_risk_70 = sum(attrition_risk_score >= 0.7),
    medium_risk_40_70 = sum(attrition_risk_score >= 0.4 & attrition_risk_score < 0.7),
    low_risk_40 = sum(attrition_risk_score < 0.4),
    total_cost_at_risk = sum(attrition_risk_score >= 0.7) * 75000
  )

risk_analysis
```

### Skills Analysis

```{r}
# Analyze skills distribution
skills_analysis <- sample_data$employees %>%
  filter(!left_company) %>%
  select(department, skills) %>%
  tidyr::separate_rows(skills, sep = ", ") %>%
  count(skills, sort = TRUE) %>%
  head(10)

skills_analysis
```

## Next Steps

1.  **Explore the Dashboard**: Launch `launch_dashboard()` for interactive exploration
2.  **Read Advanced Guides**: Check out the analytics and modeling vignettes
3.  **Customize for Your Needs**: Adapt the models and analyses for your specific use case
4.  **Schedule Regular Analysis**: Set up automated reporting for ongoing insights

## Getting Help

```{r, eval=FALSE}
# Get help on specific functions
?build_attrition_model
?generate_business_insights
?launch_dashboard

# View all package functions
help(package = "GlobalTechAnalytics")
```

For more detailed documentation and examples, visit the [package website](https://github.com/Anna-KG/GlobalTechAnalytics).
